On-the-fly Algorithms for GKAT Equivalences \thanks{Identify applicable funding agency here. If none, delete this.}

### 2025-01-19 main.tex
update results

\subsection{Performance}\label{sec:performance-implementation}
Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM. \Cref{tab:benchmark} presents the results of equivalence checking based on derivatives (DRV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK). Each benchmark consists of 50 expression pairs with properties described by the benchmark's name. For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive tests (P10) and are completely random (RD). Benchmarks with the suffix EQ have expression pairs that are known to be equivalent. The time columns measure the \textit{total time} used (in seconds) for checking all expression pairs in the benchmarks. The mem columns measure the \textit{peak memory} used (in MB).

@@

\begin{table*}
\centering
\begin{tabular}{l r r r r r r}
    Benchmark & Time (DRV) & Time (TC) & Time (SK) & Mem (DRV) & Mem (TC) & Mem (SK) \\
    \hline
    E250B5P10RD    & 0.19 & 0.18 & 5.82    & 15.36  & 14.76  & 114.06  \\
    E250B5P10EQ    & 0.21 & 0.18 & 2.83    & 15.56  & 14.95  & 100.48  \\
    E500B5P50RD    & 0.23 & 0.22 & 37.28   & 16.26  & 15.49  & 524.89  \\
    E500B5P50EQ    & 0.26 & 0.21 & 14.06   & 17.97  & 15.54  & 546.91  \\
    E1000B10P100RD & 0.31 & 0.37 & Timeout & 18.36  & 21.41  & N/A     \\
    E1000B10P100EQ & 0.37 & 0.28 & 77.83   & 20.41  & 17.66  & 5822.66 \\
    E2000B20P200RD & 1.50 & 2.75 & Timeout & 239.71 & 283.45 & N/A     \\
    E2000B20P200EQ & 3.34 & 3.03 & Timeout & 107.92 & 102.46 & N/A     \\
\end{tabular}
\caption{Benchmarks}\label{tab:benchmark}
\end{table*}

@@

\begin{onecolumn}
\begin{lstlisting}[
    caption=Dead State Detection Implementation,
    label=lst:dead-state-detection
]
pub type Deriv = Vec<(BExp, Exp, u64)>;

pub struct Solver {
    // search states
    dead_states: HashSet<Exp>,
    explored: HashSet<Exp>,
    uf_table: HashMap<Exp, UnionFindNode<()>>,
    // caching
    eps_cache: LruCache<Exp, BExp>,
    deriv_cache: LruCache<Exp, Deriv>,
}

impl Solver {
    #[inline]
    pub fn known_dead(&self, exp: &Exp) -> bool {
        self.dead_states.contains(&exp)
    }

    pub fn is_dead(&mut self, gkat: &mut Gkat, exp: &Exp) -> bool {
        let mut stack = Vec::new();
        stack.push(exp.clone());
        self.explored.clear();
        while let Some(exp) = stack.pop() {
            if self.known_dead(&exp) || self.explored.contains(&exp) {
                continue;
            }
            self.explored.insert(exp.clone());
            let eps = self.epsilon(gkat, &exp);
            if eps.is_false() {
                for (b, e, _) in self.derivative(gkat, &exp) {
                    // check is not strictly needed due to eager-pruning
                    if b.is_false() {
                        continue;
                    }
                    stack.push(e);
                }
            } else {
                return false;
            }
        }
        self.dead_states.extend(self.explored.iter().cloned());
        return true;
    }
}
\end{lstlisting}
\end{onecolumn}

>>>>

\subsection{Performance}\label{sec:performance-implementation}
Our experiments are performed on a laptop with an Apple M4 Pro CPU and 24 GB RAM. \Cref{tab:benchmark-time,tab:benchmark-memory} present the results of equivalence checking based on derivatives (DV), Thompson's construction (TC) and SymKAT~\cite{pous_SymbolicAlgorithmsLanguage_2015} (SK). Each benchmark consists of 50 expression pairs with properties described by the benchmark's name. For example, expressions in E250B5P10RD have around 250 primitive actions in total (E250), a maximum boolean expression size of 5 (B5), 10 possible unique primitive tests (P10) and are completely random (RD). Benchmarks with the suffix EQ have expression pairs that are known to be equivalent. The time columns measure the \textit{total time} used (in seconds) for checking all expression pairs in the benchmarks. The mem columns measure the \textit{peak memory} used (in MB).

@@

\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10rd    &   0.19 &   0.18 & 0.20 & 0.23 &    5.82 \\
    e250b5p10eq    &   0.21 &   0.18 & 0.18 & 0.29 &    2.83 \\
    e500b5p50rd    &   0.23 &   0.22 & 0.27 & 0.39 &   37.28 \\
    e500b5p50eq    &   0.26 &   0.21 & 0.28 & 0.57 &   14.06 \\
    e1000b10p100rd &   0.31 &   0.37 & 0.48 & 0.85 & timeout \\
    e1000b10p100eq &   0.37 &   0.28 & 0.44 & 1.04 &   77.83 \\
    e2000b20p200rd &   1.50 &   2.75 & 0.99 & 2.01 & timeout \\
    e2000b20p200eq &   3.34 &   3.03 & 0.67 & 3.14 & timeout \\
    e3000b30p200rd &   1.75 &  26.38 & 1.91 & 4.46 & timeout \\
    e3000b30p200eq &  17.86 &  22.59 & 1.09 & 5.25 & timeout \\
    degenerate     & 220.53 & 232.63 & 0.42 & 0.93 & timeout
\end{tabular}
\caption{Benchmark Time Usage}\label{tab:benchmark-time}
\end{table}

\begin{table}
\footnotesize
\centering
\setlength{\tabcolsep}{3pt}
\begin{tabular}{l r r r r r}
    Benchmark & DV$_{\textsf{BDD}}$ & TC$_{\textsf{BDD}}$ & DV$_{\textsf{SAT}}$ & TC$_{\textsf{SAT}}$ & SK \\
    \hline
    e250b5p10rd    &  15.36 &   14.76 &  6.81 &  7.02 &  114.06 \\
    e250b5p10eq    &  15.56 &   14.95 &  7.06 &  7.25 &  100.48 \\
    e500b5p50rd    &  16.26 &   15.49 &  7.02 &  7.25 &  524.89 \\
    e500b5p50eq    &  17.97 &   15.54 &  7.07 &  7.27 &  546.91 \\
    e1000b10p100rd &  18.36 &   21.41 &  8.22 &  7.99 & timeout \\
    e1000b10p100eq &  20.41 &   17.66 &  8.69 &  7.63 & 5822.66 \\
    e2000b20p200rd & 239.71 &  283.45 & 13.33 & 11.92 & timeout \\
    e2000b20p200eq & 107.92 &  102.46 & 13.20 & 14.02 & timeout \\
    e3000b30p200rd & 112.59 & 1235.46 & 21.44 & 20.10 & timeout \\
    e3000b30p200eq & 245.92 &  228.85 & 21.41 & 18.15 & timeout \\
    degenerate     & 631.47 & 1229.30 & 19.62 & 18.43 & timeout
\end{tabular}
\caption{Benchmark Memory Usage}\label{tab:benchmark-memory}
\end{table}

@@

\begin{onecolumn}
\begin{lstlisting}[
    caption=Dead State Detection Implementation,
    label=lst:dead-state-detection
]
pub type Deriv<B> = Vec<(B, Exp<B>, u64)>;

pub struct Solver<B> {
    // search states
    dead_states: HashSet<Exp<B>>,
    explored: HashSet<Exp<B>>,
    uf_table: HashMap<Exp<B>, UnionFindNode<()>>,
    // caching
    eps_cache: HashMap<Exp<B>, B>,
    drv_cache: HashMap<Exp<B>, Deriv<B>>,
}

impl<B: BExp> Solver<B> {
    #[inline]
    pub fn known_dead(&self, exp: &Exp<B>) -> bool {
        self.dead_states.contains(&exp)
    }

    pub fn is_dead<G: Gkat<B>>(&mut self, gkat: &mut G, exp: &Exp<B>) -> bool {
        let mut stack = Vec::new();
        stack.push(exp.clone());
        self.explored.clear();
        while let Some(exp) = stack.pop() {
            if self.known_dead(&exp) || self.explored.contains(&exp) {
                continue;
            }
            self.explored.insert(exp.clone());
            let eps = self.epsilon(gkat, &exp);
            if gkat.is_false(&eps) {
                for (_, e, _) in self.derivative(gkat, &exp) {
                    stack.push(e);
                }
            } else {
                return false;
            }
        }
        self.dead_states.extend(self.explored.iter().cloned());
        return true;
    }
}
\end{lstlisting}
\end{onecolumn}
